---
title: "Worldwide Food Production"
output: html_notebook
---

This is a dataset of worldwide food production. 
https://www.kaggle.com/dorbicycle/world-foodfeed-production

The goal of this report will be to analyze how food production has changed over time. I am particularly interested in newly-industrialized countries, which have not yet attained developed status, but are developing very fast compared to other states.
https://en.wikipedia.org/wiki/Newly_industrialized_country

First, I will produce a world map of how worldwide wheat production has changed over time.

```{r echo=TRUE, include=FALSE}
library(tidyverse)
library(leaflet)
library(rworldmap)

##load data
load(file = "food.rdata")

wheat <- filter(food, Item == "Wheat and products", Element == "Food")

wheat1970 <- filter(wheat, Year == 1970)
wheat1990 <- filter(wheat, Year == 1990)

```

```{r}
ggplot(dat[dat$Item=="Wheat and products",]) + geom_line(aes(x = Year, y = Amount, color = Element)) + facet_wrap(~Area)
```

normalize countries by population
explore foods with large outputs
make a map just to provide context



```{r}
##mapCountryData(matched, nameColumnToPlot="Amount", mapTitle="Worldwide Production of Wheat in 1970", catMethod = "pretty", colourPalette = "heat")

##mapCountryData(matched2, nameColumnToPlot="Amount", mapTitle="Worldwide Production of Wheat in 1990", catMethod = "pretty", colourPalette = "heat")

##match data to country
##matched <- joinCountryData2Map(wheat1970, joinCode = "ISO3", nameJoinColumn = "Area.Abbreviation")
##matched2 <- joinCountryData2Map(wheat1990, joinCode = "ISO3", nameJoinColumn = "Area.Abbreviation")

```
